Evaluating the Performance and Energy Efficiency of the COSMO-ART model system

### 2014-04-24 paper_enahpc14/Sec3:Measurement-framework.tex
Manuel's changes

\subsubsection{UHAM - UJI}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13} that works in  combination with Extrae and Paraver,
which are profiling/tracing and visualization tools, respectively.

%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To use our  approach, COSMO-ART is compiled using  the Extrae compiler
wrappers,  which  automatically instrument  the  Fortran  code of  the
model. Next, COSMO-ART  is run on the nodes,  thus dissipating certain
amount  of power.   These  nodes are  connected  to power  measurement
devices that account for the dissipated power/consumed energy and send
the power  data to the  tracing server.  The client,  meanwhile, sends
start/stop  primitives  in  order  to  gather  captured  data  by  the
watt-meters onto the  tracing server, where an instance  of the \pmlib
server is running.

Once COSMO-ART run is finished, a file containing the power profile is
created using the power data  received from the tracing server and the
instrumentation post-processing generates the performance trace files.
All  these files  are  combined then  in  Paraver which  allows us  to
visualise the performance trace and the power profile of COSMO-ART all
together.

>>>>

\subsubsection{Power-performance measurement framework at UJI}

To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the \pmlib power measurement framework~\cite{energy13} that works in  combination with Extrae and Paraver~\ref{paraver},
which are profiling/tracing and visualization tools, respectively.

To use our approach, COSMO-ART is compiled using the Extrae library, 
which  automatically instruments the Fortran  code of  the
model in order to generate traces that record, e.g., MPI calls.
Next, COSMO-ART  is run on the nodes,  thus dissipating certain
amount  of power. These nodes are connected to power measurement devices that account
for the dissipated power/consumed energy and send the power data to the tracing server. 
The client,  meanwhile, sends start/stop  primitives  in  order  to  gather  captured  data  by  these PDUs onto the  tracing server, where an instance  of the \pmlib server is running.

Once COSMO-ART run is finished, trace files containing the power profile 
using the power data  received from the tracing server and the
instrumentation trace files are generated.
Afterwards, the combination of both trace files in Paraver allows us 
to visualise and correlate the performance trace with the power profile of COSMO-ART.

### 2014-04-24 paper_enahpc14/Sec4:Evaluation.tex
Manuel's changes

\paragraph{CSCS - Monch} 
~\\
\vskip -1em
Monch is  a 10 rack  NEC-provided and dual-socket Intel  Ivy Bridge-EP
based  cluster, utilised  by  scientists  who are  part  of the  Swiss
Platform     for      Advanced     Scientific     Computing     (PASC,
\url{http://www.pasc-ch.org/}). It is composed of 312 standard compute
nodes, 24 large-memory compute nodes and 24 huge-memory compute nodes.
Each standard  compute node comprises  two Intel Ivy  Bridge Efficient
Performance (EP)  E5-2660 v2 ten-core processors operating  at 2.2 GHz
base  clock speed,  themselves connected  by a  high  speed InfiniBand
network based on Mellanox SX6036  managed FDR switches, with a 56 Gb/s
speed.  Each core  has 32 KB instruction and 32 KB  data L1 caches and
256  KB of L2  cache. All  10 cores  share a  25 MB  L3 cache  and the
platform has  32GB of  DDR3 1600 MHz  RAM. For  our energy-to-solution
benchmark, a  full rack  of Monch constituted  of 52  standard compute
nodes (monchc[029-080]) was considered.

\paragraph{CSCS - Pilatus} 
~\\
\vskip -1em
Pilatus  is  a dual-socket  eight-core  Intel  Sandy  Bridge EP  based
cluster used as Piz Daint pre-post processing cluster.  It is composed
of 42 compute  nodes and has 2 high-speed  interconnects based on FDR:
the  first is  dedicated to  the  MPI traffic  and the  second to  the
storage high  speed traffic.   The 2 login  nodes and the  42 computes
nodes  consists in  11  twin-pair Intel  E5-Series  DALCO r2264i4t  2U
scalable compute modules.  Each  module contains 4 compute nodes based
on two Intel  Xeon E5-2670 processors operating at  2.6 GHz base clock
speed, themselves  connected by a high speed  InfiniBand network based
on Mellanox  SX6036 managed FDR switches,  with a 56  Gb/s speed. Each
core has 32 KB  instruction and 32 KB data L1 caches  and 256 KB of L2
cache.  All  the 8 core share  a 20 MB  L3 cache and the  platform has
64GB of  DDR3 1600 MHz  RAM.  For our energy-to-solution  benchmark, a
full  rack  of  Pilatus  constituted  of  42  standard  compute  nodes
(pilatus[03-44]) was considered.

\paragraph{UJI - Tintorrum} 
~\\
\vskip -1em
Tintorrum  is a heterogeneous  cluster composed  of 28  compute nodes.
For our experiments only a subset of these nodes were considered. This
set is  composed of 16  nodes, each of  which includes two  Intel Xeon
E5645 hexa-core  processors at 2.40  GHz connected via  Infiniband QDR
(Mellanox MTS3600 switch).  Each core  has 32 KB instruction and 32 KB
data L1 caches and 256 KB L2 cache. The 6 cores share a 12 MB L3 cache
and the platform has 288 GB of DDR3 1333 MHz.

@@

\begin{figure}[htbf]
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-2.eps}
  \caption{Time-to-solution and energy-to-solution on Tintorrum -
    Blocking mode}
  \label{fig:3}
\end{figure}

>>>>

\begin{itemize}

\item \textsc{Monch} is a 10 rack NEC-provided and dual-socket Intel  Ivy Bridge-EP
based  cluster, utilised  by  scientists  who are  part  of the  Swiss
Platform     for      Advanced     Scientific     Computing     (PASC,
\url{http://www.pasc-ch.org/}). It is composed of 312 standard compute
nodes, 24 large-memory compute nodes and 24 huge-memory compute nodes.
Each standard  compute node comprises  two Intel Ivy  Bridge Efficient
Performance (EP)  E5-2660 v2 ten-core processors operating  at 2.2 GHz
base  clock speed,  themselves connected  by a  high  speed InfiniBand
network based on Mellanox SX6036  managed FDR switches, with a 56 Gb/s
speed.  Each core  has 32 KB instruction and 32 KB  data L1 caches and
256  KB of L2  cache. All  10 cores  share a  25 MB  L3 cache  and the
platform has  32GB of  DDR3 1600 MHz  RAM. For  our energy-to-solution
benchmark, a  full rack  of Monch constituted  of 52  standard compute
nodes (monchc[029-080]) was considered.

\item \textsc{Pilatus} is a dual-socket eight-core Intel Sandy Bridge EP based
cluster used as Piz Daint pre-post processing cluster.  It is composed
of 42 compute  nodes and has 2 high-speed  interconnects based on FDR:
the  first is  dedicated to  the  MPI traffic  and the  second to  the
storage high  speed traffic.   The 2 login  nodes and the  42 computes
nodes  consists in  11  twin-pair Intel  E5-Series  DALCO r2264i4t  2U
scalable compute modules.  Each  module contains 4 compute nodes based
on two Intel  Xeon E5-2670 processors operating at  2.6 GHz base clock
speed, themselves  connected by a high speed  InfiniBand network based
on Mellanox  SX6036 managed FDR switches,  with a 56  Gb/s speed. Each
core has 32 KB  instruction and 32 KB data L1 caches  and 256 KB of L2
cache.  All  the 8 core share  a 20 MB  L3 cache and the  platform has
64GB of  DDR3 1600 MHz  RAM.  For our energy-to-solution  benchmark, a
full  rack  of  Pilatus  constituted  of  42  standard  compute  nodes
(pilatus[03-44]) was considered.

\item \textsc{Tintorrum} is an heterogeneous  cluster composed  of 28  compute nodes.
For our experiments only a subset of these nodes were considered. This
set is  composed of 16  nodes, each of  which includes two  Intel Xeon
E5645 hexa-core  processors at 2.40  GHz connected via  Infiniband QDR
(Mellanox MTS3600 switch).  Each core  has 32 KB instruction and 32 KB
data L1 caches and 256 KB L2 cache. The 6 cores share a 12 MB L3 cache
and the platform has 288 GB of DDR3 1333 MHz.

\end{itemize}

@@

\begin{figure*}[htbf]
  \centering
  \scriptsize
  MPI polling mode\\
  \includegraphics[width=0.78\textwidth]{Figs/23_24_blq1_stat.eps}\\
  MPI blocking mode\\
  \includegraphics[width=0.78\textwidth]{Figs/23_24_blq0_stat.eps}\\
  \caption{blabla}
  \label{fig:3}
\end{figure*}

\begin{figure*}[htbf]
  \centering\hspace{0.5cm}
  \scalebox{0.5}{\input{Figs/24hour.pstex_t}}
  \caption{24 hours simulation trace using the MPI blocking mode on Tintorrum.}
  \label{fig:3}
\end{figure*}

\begin{figure*}[htbf]
  \centering
  \scriptsize
  Time frame from 11\,h to 12\,h using the MPI polling mode\\
  \scalebox{0.5}{\input{Figs/11_12_blq0_figure.pstex_t}}\\
  Time frame from 11\,h to 12\,h using the MPI blocking mode
  \scalebox{0.5}{\input{Figs/11_12_blq1_figure.pstex_t}}\\
  Time frame from 23\,h to 24\,h using the MPI polling mode
  \scalebox{0.5}{\input{Figs/23_24_blq0_figure.pstex_t}}\\
  Time frame from 23\,h to 24\,h using the MPI blocking mode\\
  \scalebox{0.5}{\input{Figs/23_24_blq1_figure.pstex_t}}
  \begin{tabular}{rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rlp{-0.2cm}rl}
& &  & &  & &  & &  \\[-0.15cm]
Computation     & \multicolumn{1}{>{\columncolor[RGB]{  0,  , 255}}p{0.4cm}}{}  & & 
Waiting a msg.  & \multicolumn{1}{>{\columncolor[RGB]{255,  0  ,0}}p{0.4cm}}{}  & & 
Block. send     & \multicolumn{1}{>{\columncolor[RGB]{255,  0,174}}p{0.4cm}}{}  & & 
Synchronization & \multicolumn{1}{>{\columncolor[RGB]{179,  0,  0}}p{0.4cm}}{}  & & 
Wait/Wait all   & \multicolumn{1}{>{\columncolor[RGB]{235,  0,  0}}p{0.4cm}}{}  \\
& &  & &  & &  & &  \\[-0.15cm]
Immediate recv. & \multicolumn{1}{>{\columncolor[RGB]{100,100,177}}p{0.4cm}}{}  & &
I/O             & \multicolumn{1}{>{\columncolor[RGB]{172,174, 41}}p{0.4cm}}{}  & &
Group comm.     & \multicolumn{1}{>{\columncolor[RGB]{255,144, 26}}p{0.4cm}}{}  & &
Others          & \multicolumn{1}{>{\columncolor[RGB]{192,224,  0}}p{0.4cm}}{}  & &
Idle            & \multicolumn{1}{>{\columncolor[RGB]{117,195,255}}p{0.4cm}}{}  \\
\end{tabular}
  \caption{Traces for 1 hour simulation of COSMO-ART leveraging the polling and blocking MPI policies on Tintorrum.}
  \label{fig:3}
\end{figure*}
